An event-sourced and distributed database with IPLD and materialized views

### 2019-09-20 draft.tex
cite: adds multiple citations, a bib, and more

When developing large-scale software systems, it is common to store data in a relational database management system (RDBMS). To model realistic systems, this type of framework often requires complex techniques for mapping data between domain models and database tables, where the same data model is used to both query and update a database. An powerful alternative is to use a set of append only logs to model the state of an object simply by applying its change sequence in the correct order. This concept can be expressed succinctly by the state machine approach [ref]: if two identical, deterministic processes begin in the same state and get the same inputs in the same order, they will produce the same output and end in the same state. This is a powerful concept baked into a simple structure, and is at the heart of many distributed database systems [ref].

@@

In modern applications it is critical to have reliable mechanisms for publishing updates and events (i.e., to support event-driven architectures), scalability (optimized write and read operations), forward-compatible application updates (e.g., code changes, retroactive events, and others), auditing systems, etc. To support such requirements, developers have begun to utilize event sourcing and command query responsibility segregation (CQRS) patterns [ref], relying on append only logs to support immutable state histories. Indeed, a number of commercial and open source software projects have emerged in recent years that facilitate event sourcing and CQRS-based applications, including Event Store [ref], Apache Kafaka [ref] and Samza [ref], among others [see ref].

@@

A key principal of ES and append only logs is that all changes to application state are stored as a sequence of events. Because any given state is simply the result of a series of atomic updates, the log can be used to reconstruct past states or process retroactive updates [ref]. The same principal means a log can be viewed as a mechanism to support an infinite number of valid state interpretations (see section  \ref{sec:ViewsProjections}). In other words, with minimal conformity, a single log can model multiple application states [ref].

@@

\begin{definition} (Projection). An event handler and corresponding reducer/fold function used to build and maintain a view from a set of (filtered) events. While projections may lead to the generation of new events, their reducer should be a pure function (see [ref] and [ref] for examples from existing systems). \end{definition}

@@

Views themselves are enabled by projections. Projections can be thought of as transformations or reducers that are applied to each event in a stream of events. They update the data backing the views, be this in memory, or persisted to a database. In a distributed setting, it may be necessary for projections to define and operate as eventually consistent data structures, to ensure all peers operating on the same stream of events have a consistent representation of the data.

@@

The CAP theorem [ref, ref] states that a distributed database can guarantee only two of the following three promises at the same time: consistency (i.e., that every read receives the most recent write or an error), availability (i.e., that every request receives a (possibly out-of-date) non-error response, and partition tolerance (i.e., that the system continues to operate despite an arbitrary number of messages being dropped (or delayed) by the network). As such, many distributed systems are now designed to provide availability and partition tolerance by trading consistency for eventual consistency. Eventual consistency allows state replicas to diverge temporarily, but eventually arrive back to the same state again. While an active area of research, designing systems with provable eventual consistency guarantees remains challenging [ref needed].

@@

How a system provides eventual consistency is often decided based on the intended use of the system. Two well-documented categories of solutions include logs (see Definition  \ref{def:Logs}), sequences of deterministic changes, and CRDTs (see Definition section  \ref{def:CRDT}). Logs work best when there is only a single-writer or complete event causality is known. In many distributed systems with multiple log writers, a minimum requirement for synchronization through logs is that the essential order of events is respected and can be determined [3, 8]. For these cases, logical clocks are a useful tool for eventual consistency and total ordering [ref]. However, some scenarios (e.g., temporarily missing events, or ambiguous order) can force a replica into a state that cannot be later resolved without costly recalculation. In specific cases, CRDTs can provide an alternative to log-based consensus (see below).

@@

CRDTs (see Definition \ref{def:CRDT}) are one way to achieve strong eventual consistency, where once all replicas have received the same events, they will arrive at the same final state, regardless of ordering. A review of the types of possible CRDTs is beyond the scope of this paper, however, it is important to note their role in eventually consistent systems and how they relate to clock-based event ordering. See for example [Merkle Clock ref] and [ref] for informative reviews of these types of data structures.

@@

Internet application architecture today is often designed as a system of clients (users) communicating to endpoints (hosts). Communication between clients and endpoints (servers) happens via the TCP/IP protocol stack and therefore largely relies on a mechanism referred to as location-based addressing. Location-based addressing, where the client makes a request that is routed to a specific endpoint based on prior knowledge (e.g., the domain name or IP address), works relatively well for many use-cases. However, there are many reasons why addressing content by location is problematic, such as duplication of storage, inefficient use of bandwidth, link rot, centralized control, and authentication issues. An alternative to location addressing, called content addressing may provide a solution to those problems. Content addressing is where the content itself is used to create an address and retrieve that content from a network [2].

@@

There are a number of systems that utilize content addressing to access content and information [list some as refs], and there is an active body of literature covering its design and implementation [ref]. The Interplanetary File System (IPFS) — which is a set of protocols to create a content-addressed, peer-to-peer filesystem [1] — is one such system. In IPFS, the address for any content is determined based on a cryptographic hash of the content itself. In practice, the IPFS Content IDentifier (CID) is a multihash, which is a self-describing “protocol for differentiating outputs from various well-established cryptographic hash functions, addressing size + encoding considerations” [ref]. That addressing system confers several benefits to the network, including tamper resistance (we can be confident that a given piece of content has not been modified en route if its hash matches what we were expected/requested) and de-duplication (because the same content from different peers will produce the same hash address). Additionally, IPFS content addresses are immutable and universally unique.

@@

While content addressing doesn’t tell you how to get a file, IPFS (via libp2p [ref]) provides a system for moving files across the network.  On the IPFS network, a client who wants specific content requests the CID from the network of IPFS hosts. The client's request is routed to the first host capable of fulfilling the request — i.e., the first host that is actively storing the content behind the given CID. The IPFS network can be seen as a distributed file system, with all of the benefits [ref] that come with this type of file system design.

@@

\section{The Threads protocol}

@@

\begin{example*}
\centering
\begin{minipage}{0.7\textwidth}
\begin{lstlisting}[language=json,firstnumber=1]
# Log multiaddress
 /ipel/12D3KooWC2zyCVron7AA34N6oKNtaXaZB51feG9rBkr7QbCcW8ab

# Encapsulated multiaddress
 /ip4/127.0.0.1/tcp/1234/p2p/12D..dwaA6Qe/ipel/12D..bCcW8ab

# Address book
[
  /p2p/12D..dwaA6Qe/ipel/12D..bCcW8ab,
  /p2p/12D..dJT6nXY/ipel/12D..bCcW8ab # Follower
]
\end{lstlisting}
\end{minipage}
  \caption{The Log Multiaddress.}
  \label{ex:Multiaddress}
\end{example*}

@@

\begin{example}
\begin{lstlisting}
"bafyoiobghzefwlidfrwkqmzz2ka66zgmdmgeobw2mimktr5jivsavya"
\end{lstlisting}
\caption{V1, 256 bit.}
\end{example} \label{ex:v1256}

@@

\section{Latex Examples}

\begin{figure}
  \includegraphics[width=\linewidth]{boat.jpg}
  \caption{A boat.}
  \label{fig:boat1}
\end{figure}

\begin{quotation}
A quote
\end{quotation}

\begin{figure*}
  \includegraphics[width=\textwidth,height=4cm]{boat.jpg}
  \caption{This is a full width boat.}
  \label{fig:boat1}
\end{figure*}

Here is a citation models~\cite{gelenbe06}. The approach constructs

here is a ref provides performance measures of interest to the SB and to the
seller, is discussed in Section \ref{Performance} where we first
discuss how the SB can behave in order to optimise outcomes that
are in its best interest, and provide  numerical examples to
illustrate the approach and the model predictions. We then explore
how the SB can try to achieve balance and compete with the other
bidders in Section \ref{KeepUp}. Finally Section
\ref{Price-Dependent} generalises the analysis to the case where
the bidding rates depend on the current price attained in the
auction. Conclusions are drawn in Section \ref{Conclusions} where
we also suggest further work.

@@

\bibliographystyle{compj}
% \bibliography{ModellingBidders}
\input{whitepaper.bbl}

>>>>

When developing large-scale software systems, it is common to store data in a relational database management system (RDBMS). To model realistic systems, this type of framework often requires complex techniques for mapping data between domain models and database tables, where the same data model is used to both query and update a database. A powerful alternative is to use a set of append only logs to model the state of an object simply by applying its change sequence in the correct order. This concept can be expressed succinctly by the state machine approach \cite{schneiderImplementingFaulttolerantServices1990}: if two identical, deterministic processes begin in the same state and get the same inputs in the same order, they will produce the same output and end in the same state. This is a powerful concept baked into a simple structure, and is at the heart of many distributed database systems \cite{LogWhatEvery}.

@@

In modern applications it is critical to have reliable mechanisms for publishing updates and events (i.e., to support event-driven architectures), scalability (optimized write and read operations), forward-compatible application updates (e.g., code changes, retroactive events, and others), auditing systems, etc. To support such requirements, developers have begun to utilize event sourcing and command query responsibility segregation (CQRS) patterns \cite{bettsExploringCQRSEvent2013}, relying on append only logs to support immutable state histories. Indeed, a number of commercial and open source software projects have emerged in recent years that facilitate event sourcing and CQRS-based applications, including Event Store \cite{EventStore}, Apache Kafaka \cite{ApacheKafka} and Samza \cite{Samza}, among others (see \cite{kleppmannDesigningDataintensiveApplications2017}).

@@

A key principal of ES and append only logs is that all changes to application state are stored as a sequence of events. Because any given state is simply the result of a series of atomic updates, the log can be used to reconstruct past states or process retroactive updates \cite{fowlerEventSourcing}. The same principal means a log can be viewed as a mechanism to support an infinite number of valid state interpretations (see section  \ref{sec:ViewsProjections}). In other words, with minimal conformity, a single log can model multiple application states \cite{microsoftcorporationAzureApplicationArchitecture}.

@@

\begin{definition} (Projection). An event handler and corresponding reducer/fold function used to build and maintain a view from a set of (filtered) events. While projections may lead to the generation of new events, their reducer should be a pure function (see \cite{eventsouceEventSourcingBasics} and \cite{redux.js.orgGettingStartedRedux} for examples from existing systems). \end{definition}

@@

Views themselves are enabled by projections\footnote{Terminology in this section may differ from some other examples of ES and CQRS patterns, but reflects the underlying architecture and designs that the Textile team will elaborate on in \ref{sec:ThreadsProtocol}}. Projections can be thought of as transformations or reducers that are applied to each event in a stream of events. They update the data backing the views, be this in memory, or persisted to a database. In a distributed setting, it may be necessary for projections to define and operate as eventually consistent data structures, to ensure all peers operating on the same stream of events have a consistent representation of the data.

@@

The CAP theorem \cite{brewerRobustRobustDistributed2000,gilbertBrewerConjectureFeasibility2002a} states that a distributed database can guarantee only two of the following three promises at the same time: consistency (i.e., that every read receives the most recent write or an error), availability (i.e., that every request receives a (possibly out-of-date) non-error response, and partition tolerance (i.e., that the system continues to operate despite an arbitrary number of messages being dropped (or delayed) by the network). As such, many distributed systems are now designed to provide availability and partition tolerance by trading consistency for eventual consistency. Eventual consistency allows state replicas to diverge temporarily, but eventually arrive back to the same state again. While an active area of research, designing systems with provable eventual consistency guarantees remains challenging \cite{shapiroComprehensiveStudyConvergent2011a,almeidaDeltaStateReplicated2018}.

@@

How a system provides eventual consistency is often decided based on the intended use of the system. Two well-documented categories of solutions include logs (see Definition  \ref{def:Logs}), sequences of deterministic changes, and CRDTs (see Definition section  \ref{def:CRDT}). Logs work best when there is only a single-writer or complete event causality is known. In many distributed systems with multiple log writers, a minimum requirement for synchronization through logs is that the essential order of events is respected and can be determined \cite{schwarzDetectingCausalRelationships1994,katzInterleavingSetTemporal1990}. For these cases, logical clocks are a useful tool for eventual consistency and total ordering \cite{kulkarniLogicalPhysicalClocks2014}. However, some scenarios (e.g., temporarily missing events, or ambiguous order) can force a replica into a state that cannot be later resolved without costly recalculation. In specific cases, CRDTs can provide an alternative to log-based consensus (see below).

@@

CRDTs (see Definition \ref{def:CRDT}) are one way to achieve strong eventual consistency, where once all replicas have received the same events, they will arrive at the same final state, \emph{regardless of ordering}\footnote{Though non-commutative CRDTs may require a specific ordering of events in certain cases \cite{sanjuanMerkleCRDTs2019}}. A review of the types of possible CRDTs is beyond the scope of this paper, however, it is important to note their role in eventually consistent systems and how they relate to clock-based event ordering. See for example \cite{sanjuanMerkleCRDTs2019} and [MISSING] for informative reviews of these types of data structures.

@@

Internet application architecture today is often designed as a system of clients (users) communicating to endpoints (hosts). Communication between clients and endpoints (servers) happens via the TCP/IP protocol stack and therefore largely relies on a mechanism referred to as location-based addressing. Location-based addressing, where the client makes a request that is routed to a specific endpoint based on prior knowledge (e.g., the domain name or IP address), works relatively well for many use-cases. However, there are many reasons why addressing content by location is problematic, such as duplication of storage, inefficient use of bandwidth, link rot, centralized control, and authentication issues. An alternative to location addressing, called content addressing may provide a solution to those problems. Content addressing is where the content itself is used to create an address and retrieve that content from a network \cite{mortContentBasedAddressing2012}.

@@

There are a number of systems that utilize content addressing to access content and information [list some as refs], and there is an active body of literature covering its design and implementation [ref]. The Interplanetary File System (IPFS) — which is a set of protocols to create a content-addressed, peer-to-peer filesystem \cite{benetIpfscontentAddressedVersioned2014} is one such system. In IPFS, the address for any content is determined based on a cryptographic hash of the content itself. In practice, the IPFS Content IDentifier (CID) is a multihash, which is a self-describing “protocol for differentiating outputs from various well-established cryptographic hash functions, addressing size + encoding considerations\footnote{https://github.com/multiformats/multihash}. That addressing system confers several benefits to the network, including tamper resistance (we can be confident that a given piece of content has not been modified en route if its hash matches what we were expected/requested) and de-duplication (because the same content from different peers will produce the same hash address). Additionally, IPFS content addresses are immutable and universally unique.

@@

While content addressing doesn't tell you how to get a file, IPFS (via libp2p\footnote{https://libp2p.io/}) provides a system for moving files across the network.  On the IPFS network, a client who wants specific content requests the CID from the network of IPFS hosts. The client's request is routed to the first host capable of fulfilling the request (i.e., the first host that is actively storing the content behind the given CID). The IPFS network can be seen as a distributed file system, with all of the benefits that come with this type of file system design.

@@

\section{The Threads protocol}\label{sec:ThreadsProtocol}

@@

\begin{example*}
\centering
\begin{minipage}{0.7\textwidth}
\begin{lstlisting}
// Log multiaddress
 "/ipel/12D3KooWC2zyCVron7AA34N6oKNtaXaZB51feG9rBkr7QbCcW8ab"

// Encapsulated multiaddress
 "/ip4/127.0.0.1/tcp/1234/p2p/12D..dwaA6Qe/ipel/12D..bCcW8ab"

// Address book
[
  "/p2p/12D..dwaA6Qe/ipel/12D..bCcW8ab",
  "/p2p/12D..dJT6nXY/ipel/12D..bCcW8ab" // Follower
]
\end{lstlisting}
\end{minipage}
  \caption{The Log Multiaddress.}
  \label{ex:Multiaddress}
\end{example*}

@@

Example (V1, 256 bit): `bafyoiobghzefwlidfrwkqmzz2ka66zgmdmgeobw2mimktr5jivsavya`

@@

@@

\bibliographystyle{compj}
\bibliography{whitepaper}
